---
title: "Penguin lab - flipper lengths, ANOVA"
author: "An Bui"
date: "07 November 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### 1. Set up.
```{r set up code, message = FALSE}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")
```

#### 2. Introductory data wrangling.
**Question: is there a significant difference in mean flipper lengths for female Chinstrap, Gentoo, and Adelie penguins near Palmer Archepelago?**  
One factor: species (penguins)  
Levels: 3 (Chinstrap, Gentoo, Adelie)  
Continuous random variable: flipper length  
This means you can use a **one-way ANOVA.**  

Assume:  
- independent observations  
- equal variance (Levene's test)  
- normality (histograms/qqplots)  

First, count up your data.
```{r count data using with()}
# with dataset penguins, create a table of counts
counts <- with(penguins, table(Species, Sex))
```
Seems like there are >30 samples, so you can use CLT to justify an ANOVA.

```{r}
# create a new data frame from penguins
penguins2 <- penguins %>% 
  # add a new column using mutate
  mutate(
    # call that column common_name
    common_name = 
      # use case_when for common names of penguins to match with the other column of both scientific and common
      case_when(
        Species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
        Species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
        Species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
      )
  ) %>% 
  # retain all columns except Species
  select(-Species) %>% 
  # filter for female penguins
  filter(Sex == "FEMALE")
```




