---
title: "Penguin lab - flipper lengths, ANOVA"
author: "An Bui"
date: "07 November 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### 1. Set up.
```{r set up code, message = FALSE}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")
```

#### 2. Introductory data wrangling.
**Question: is there a significant difference in mean flipper lengths for female Chinstrap, Gentoo, and Adelie penguins near Palmer Archepelago?**  
One factor: species (penguins)  
Levels: 3 (Chinstrap, Gentoo, Adelie)  
Continuous random variable: flipper length  
This means you can use a **one-way ANOVA.**  

Assume:  
- independent observations  
- equal variance (Levene's test)  
- normality (histograms/qqplots)  

Figure out of your data fulfill these assumptions. First, count up your data.
```{r count data using with()}
# with dataset penguins, create a table of counts
counts <- with(penguins, table(Species, Sex))
```
Seems like there are >30 samples, so you can use CLT to justify an ANOVA.

```{r simplify data frame to use}
# create a new data frame from penguins
penguins2 <- penguins %>% 
  # add a new column using mutate
  mutate(
    # call that column common_name
    common_name = 
      # use case_when for common names of penguins to match with the other column of both scientific and common
      case_when(
        Species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
        Species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
        Species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
      )
  ) %>% 
  # retain all columns except Species
  select(-Species) %>% 
  # filter for female penguins
  filter(Sex == "FEMALE")
```

Now, make some exploratory graphs.
```{r histogram}
penguins_hist <- ggplot(penguins2, aes(x = flipper_length)) +
  geom_histogram(aes(fill = common_name),
                 bins = 8) +
  facet_wrap (~ common_name) +
  theme_bw() +
  labs(x = "Flipper length (mm)", y = "Frequency (n = ?)")

penguins_hist
```

```{r penguins qqplot}
penguins_qq <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq() +
  facet_wrap(~ common_name) +
  theme_bw()

penguins_qq
```
Based on qqplots and histograms, observations for each group seems normally distributed. Also, n > 30 so means will be normally distributed by CLT.  

Then, test for equal variances. If the largest sample variance is < 4x greater than the samllest sample variance, then these are "close enough" to consider variances equal.
```{r test for variances}
# new data frame from penguins
variances <- penguins2 %>% 
  # group by species
  group_by(common_name) %>% 
  summarize(
    variance = var(flipper_length)
  )
```
Largest variance (33) is not more than 4x greater than the smallest variance (15).

Then, use a Levene's test (in `car` package).  
**H~0~:** No difference in variances.  
**H~A~:** Variances are _not_ equal.
```{r Levene test}
# do a Levene's test
levene_penguin <- leveneTest(flipper_length ~ common_name, data = penguins2)

levene_penguin
```
There is no significant difference in variances across species. Therefore, you can do a one-way ANOVA.

#### 3. ANOVA and post-hoc testing
**H~0~:** All means are the same.
**H~A~:** At least two means are different.
```{r ANOVA}
# compare means of random variable flipper_length based on common_name
penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)

# use summary() to get outcome of aov()
summary(penguin_aov)
```
So, at least two means are different. But which ones are actually different? Use a post-hoc Tukey's HSD.
```{r}
# do a Tukey's HSD on your ANOVA
penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph
```
In every pairwise comparison, means are different.  
**Statement:** Mean flipper lengths of female penguins differ signficantly for all species studied (one-way ANOVA, F(df) = statistic, p < 0.001, $\alpha$ = 0.05 with post-hoc Tukey's HSD, $\alpha$ = 0.05).  
`aov.test()` function in `onewaytests` package allows inline references, not `aov()`.


